A Speculative Approach to Parallelization in Particle Swarm Optimization

### 2011-02-10 honors_thesis/si_paper/reviewer_response/response.tex

Response: We added the citation after the first sentence of the paper.

\subsection*{1.8}

Reviewer comments: What serves as the motivation for the authors's work is
insufficiently discussed.  The authors say that adding more particles leads to
``diminishing returns''. This may be true when dealing with constant size
populations.  However, a recent trend is to vary the population size over time.
In fact, it has been shown that adding particles \emph{over time} may actually
solve the problem the authors refer to. Some references are ``Incremental
Social Learning in Particle Swarms'' by Montes de Oca and others published in
IEEE SMCB and ``Efficient Population Utilization Strategy for Particle Swarm
Optimizer'' by Hsieh and others in the same journal.

Response: We added a paragraph in our related work section discussing these
methods (the last paragraph of the section).  We also added a paragraph to
future work discussing how these methods are particularly amenable to
speculative evaluation when parallelized (the second paragraph of the section).

@@

\subsection*{1.9}

Reviewer comments: In the first paragraph of Section 3, the authors say that
adding particles may not help past a certain swarm size. This is true if the
computation time allocated to the optimization process remains fixed. However,
in some problems, it is the solution quality what matters. In that case, adding
particles \emph{and} computational budget may indeed be the best thing to do.

Response: This is very much related to the previous item, and so the response
to that item is relevant here.  We also added some text to to the introduction
(paragraph beginning with ``In this paper we consider\ldots'') to better
explain our intent; that is, given a fixed computational budget, how can
processors be used most efficiently?  The text we added in this and the
previous item addresses this question.

### 2011-02-10 masters_thesis/si_paper/paper.tex
Moved the si_paper to masters_thesis, because that's what it is now

At each iteration of constricted PSO, the position $\ppos_t$ and velocity
$\pvel_t$ of each particle are updated as follows:
\begin{align}
\nonumber
	\pvel_{t+1} &=
		\constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{t}\otimes(\pbest_{t} - \ppos_{t}) \\
\label{eq:velupdate}
			& \quad \quad \quad \, + \ncoeff\nURand_{t}\otimes(\nbest_{t} - \ppos_{t})
		\bigr] \\
\label{eq:posupdate}
	\ppos_{t+1} &= \ppos_{t} + \pvel_{t+1}
\end{align}
where \( \pURand_{t} \) and \( \nURand_{t} \) are vectors of independent random
numbers drawn from a standard uniform distribution, the \( \otimes \) operator
is an element-wise vector multiplication, $\pbest$ (called personal best) is
the best position the current particle has seen, and $\nbest$ (called
neighborhood best) is the best position the neighbors of the current particle
have seen~\citep{bratton-2007-defining-a-standard-for-pso}.  The parameters \(
\ncoeff \), \( \pcoeff \), and \( \constriction \) are given prescribed values
required to ensure convergence (2.05, 2.05, and .73,
respectively)~\citep{clerc-2002-constricted-pso}.

@@

With regard to increasing the swarm size in PSO, some recent work has suggested
that increasing the swarm size through the course of the optimization process
provides better results than having a set swarm
size~\citep{hsieh-2009-efficient-population-utilization-for-pso,
montes-de-oca-2010-incremental-social-learning-pso}.  However, these results
are based on total number of function evaluations, which, when running in
parallel on expensive functions, is less important than total number of
iterations.  Other work focusing on parallelization has shown that when extra
processors are available they should be used, as performance increases with
swarm size when measuring in terms of number of
iterations~\cite{mcnabb-2009-large-particle-swarms,
jin-2005-pso-antenna-designs}.  If the swarm size were varied throughout the
course of the optimization process, some processors would be sitting idle at
most iterations.

>>>>

At each iteration of constricted PSO, the position $\ppos_t$ and velocity
$\pvel_t$ of each particle are updated as follows:
\begin{align}
\label{eq:velupdate}
	\pvel_{t+1} &=
		\constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{t}\otimes(\pbest_{t} - \ppos_{t}) +
			\ncoeff\nURand_{t}\otimes(\nbest_{t} - \ppos_{t})
		\bigr] \\
\label{eq:posupdate}
	\ppos_{t+1} &= \ppos_{t} + \pvel_{t+1}
\end{align}
where \( \pURand_{t} \) and \( \nURand_{t} \) are vectors of independent random
numbers drawn from a standard uniform distribution, the \( \otimes \) operator
is an element-wise vector multiplication, $\pbest$ (called personal best) is
the best position the current particle has seen, and $\nbest$ (called
neighborhood best) is the best position the neighbors of the current particle
have seen~\citep{bratton-2007-defining-a-standard-for-pso}.  The parameters \(
\ncoeff \), \( \pcoeff \), and \( \constriction \) are given prescribed values
required to ensure convergence (2.05, 2.05, and .73,
respectively)~\citep{clerc-2002-constricted-pso}.

@@

With regard to increasing the swarm size in PSO, some recent work has suggested
that increasing the swarm size through the course of the optimization process
provides better results than having a set swarm
size~\citep{hsieh-2009-efficient-population-utilization-for-pso,
montes-de-oca-2010-incremental-social-learning-pso}.  However, these results
are based on total number of function evaluations, which, when running in
parallel on expensive functions, is less important than total number of
iterations.  Other work focusing on parallelization has shown that when extra
processors are available they should be used, as performance increases with
swarm size when measuring in terms of number of
iterations~\citep{mcnabb-2009-large-particle-swarms,
jin-2005-pso-antenna-designs}.  If the swarm size were varied throughout the
course of the optimization process, some processors would be sitting idle at
most iterations.
